---
title: "BER_models"
author: "TXL"
#output: rmarkdown::github_document
#output: pdf_document
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(gridExtra)
# Make sure folder is correct
folder_name = "/Users/txl/TXL_R/Final Data"
dt = fread(paste(folder_name, "/2018dataClean.csv", sep = ""))
# Rename "Runs Scored till end of inning"
dt %>% setnames(old = c("Position Before", "Position After",
                        "Runs Scored till end of inning", "Runs_Scored", 
                        "League", "Park_Factor", "FIP_Pitcher_Control"), 
                new = c("position_before", "position_after", 
                        "runs_scored_EOI", "runs_scored", 
                        "league", "park_factor", "FIP"))


```


# Exploration of different models for BER calculation

Load 2018 Data
```{r}
# Load data is hidden for privacy
head(dt, 20)

# Make position_before and position_after factors
dt$position_before = as.factor(dt$position_before)
dt$position_after = as.factor(dt$position_after)

# Make league a factor
dt$league = as.factor(dt$league)

# Divide park_factor by 100 to turn to a percent measure
dt$park_factor = dt$park_factor / 100

# Turn FIP into numeric
dt$FIP = as.numeric(dt$FIP)

# Calculate max_runs possible scored
dt$max_runs = as.numeric(dt$first_runner != "") + 
  as.numeric(dt$second_runner != "") +
  as.numeric(dt$third_runner != "") + 1
  
```

# Preliminary Visualizations for Factors
```{r, warning=F}
## Explore park_factor
parks_dt = unique(dt[, c("Stadium", "park_factor")])
# Bargraph
ggplot(parks_dt, aes(x = Stadium, y = park_factor)) + 
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90)) + 
  ggtitle("Park Factors")
  
## Explore FIP
fip_dt = unique(dt[, c("res_pitcher", "FIP")])
# Histogram
ggplot(fip_dt, aes(x = FIP)) + 
  geom_histogram(binwidth = .5) + 
  ggtitle("Histogram of FIP")
```

Some notes here:
+ Maybe we need to remove some outliers in FIP
+ Leaving park_factor as is since it looks like a percentage measure



# Model 1: position_before only
```{r}
m1 = lm(runs_scored_EOI ~ position_before - 1, 
        dat = dt)
summary(m1)
m1_vt = m1$coefficients %>% data.table(factor = names(.), value = .)

#Graph coefficients
m1_vt %>% 
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("M1 Coefficients")

```

# Model 2: 
## Model 2 Version 1: position_before and league (no interaction)
```{r}
m2_v1 = lm(runs_scored_EOI ~ position_before + league - 1, 
        dat = dt)
summary(m2_v1)
m2_v1_vt = m2_v1$coefficients %>% data.table(factor = names(.), value = .)

m2_v1_vt %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90)) + 
  ggtitle("m2_v1 Coefficients (all factors)")
```

## Model 2 Version 2: position_before and league (include main league effect)
```{r}
m2_v2 = lm(runs_scored_EOI ~ position_before * league - 1, 
           dat = dt)
summary(m2_v2)
m2_v2_vt = m2_v2$coefficients %>% data.table(factor = names(.), value = .)

# Graph all factor values
m2_v2_vt %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m2_v2 Coefficients (all factors)")

# Graph only position_before values
m2_v2_vt[1:24, ] %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m2_v2 Coefficients (position_before only)")

# Graph interaction effect values
m2_v2_vt[25:48, ] %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m2_v2 Coefficients (interaction only)")

```

## Model 2 Version 3: position_before and league (without main league effect)
```{r}
m2_v3 = lm(runs_scored_EOI ~ PositionBefore_League - 1, 
           dat = dt)
summary(m2_v3)
m2_v3_vt = m2_v3$coefficients %>% data.table(factor = names(.), value = .)



# Graph all factor values
m2_v3_vt %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m2_v3 Coefficients (all factors)")

# Graph only position_before values
m2_v3_vt[1:24, ] %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m2_v3 Coefficients (position_before only)")

# Graph interaction effect values
m2_v3_vt[25:48, ] %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m2_v3 Coefficients (interaction only)")

```


# Model 3
## Model 3 Version 1: position_before and park_factor (no interaction)
```{r}
m3_v1 = lm(runs_scored_EOI ~ position_before + park_factor - 1, 
           dat = dt)
summary(m2_v1)
m3_v1_vt = m3_v1$coefficients %>% data.table(factor = names(.), value = .)

# Graph all factor values
m3_v1_vt %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m3_v1 Coefficients (all factors)")
```


## Model 3 Version 2: position_before and park_factor (include main park_factor effect)
```{r}
m3_v2 = lm(runs_scored_EOI ~ position_before * park_factor - 1, 
           dat = dt)
summary(m3_v2)
m3_v2_vt = m3_v2$coefficients %>% data.table(factor = names(.), value = .)

# Graph all factor values
m3_v2_vt %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m3_v2 Coefficients (all factors)")

# Graph only position_before values
m3_v2_vt[1:24, ] %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m3_v2 Coefficients (position_before only)")

# Graph interaction effect values
m3_v2_vt[25:48, ] %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m3_v2 Coefficients (interaction only)")

```


## Model 3 Version 3: position_before and park_factor (without main park_factor effect)
```{r}
m3_v3 = lm(runs_scored_EOI ~ position_before + position_before:park_factor - 1, 
           dat = dt)
summary(m3_v3)
m3_v3_vt = m3_v3$coefficients %>% data.table(factor = names(.), value = .)

# Graph all factor values
m3_v3_vt %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m3_v3 Coefficients (all factors)")

# Graph only position_before values
m3_v3_vt[1:24, ] %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m3_v3 Coefficients (position_before only)")

# Graph interaction effect values
m3_v3_vt[25:48, ] %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m3_v3 Coefficients (interaction only)")

```

# Model 4
## Model 4 Version 1: position_before and FIP (no interaction)
```{r}
m4_v1 = lm(runs_scored_EOI ~ position_before + FIP - 1, 
           dat = dt)
summary(m4_v1)
m4_v1_vt = m4_v1$coefficients %>% data.table(factor = names(.), value = .)

# Graph all factor values
m4_v1_vt %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m4_v1 Coefficients (all factors)")

```

## Model 4 Version 2: position_before and FIP (include main FIP factor)
```{r}
m4_v2 = lm(runs_scored_EOI ~ position_before * FIP - 1, 
           dat = dt)
summary(m4_v2)
m4_v2_vt = m4_v2$coefficients %>% data.table(factor = names(.), value = .)

# Graph all factor values
m4_v2_vt %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m4_v2 Coefficients (all factors)")

# Graph only position_before values
m4_v2_vt[1:24, ] %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m4_v2 Coefficients (position_before only)")

# Graph interaction effect values
m4_v2_vt[25:48, ] %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m4_v2 Coefficients (interaction only)")

```

## Model 4 Version 3: position_before and FIP (without main FIP factor)
```{r}
m4_v3 = lm(runs_scored_EOI ~ position_before + position_before:FIP - 1, 
           dat = dt)
summary(m4_v3)
m4_v3_vt = m4_v3$coefficients %>% data.table(factor = names(.), value = .)

# Graph all factor values
m4_v3_vt %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m4_v3 Coefficients (all factors)")

# Graph only position_before values
m4_v3_vt[1:24, ] %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m4_v3 Coefficients (position_before only)")

# Graph interaction effect values
m4_v3_vt[25:48, ] %>%
  ggplot(aes(x = factor, y = value)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("m4_v3 Coefficients (interaction only)")

```


# Calculating BER based on models

For each model, we need to perform the following steps for BER calculation:
1. Match each position_before with corresponding value
2. Match each position_after with corresponding value
3. Calculate value_created = runs_scored + value_pa - value_vb
4. Calculate value_max = #runnners + 1 + value_reset_same_outs
5. BER = value_created / value_max

# Model 1 BER: position_before only
```{r}
# Create a value list. Basically a dictionary for faster reference.
state = sort(unique(dt$position_before))
m1_vt$state = state
m1_vl = m1_vt$value
names(m1_vl) = m1_vt$state
m1_vl = c(m1_vl, "end inning" = 0)

# Create a condensed data.table
m1_dt = dt[, c("res_batter", "position_before", "position_after",
               "runs_scored", "max_runs", "runs_scored_EOI", "outs")]


# Calculate values of position_before
m1_dt$value_pb = sapply(m1_dt$position_before, 
                        function(pb) return(m1_vl[pb]))

# Calculate value of position_after
m1_dt$value_pa = sapply(m1_dt$position_after,
                        function(pa) return(m1_vl[pa]))

# Calculate value_max
m1_dt$value_max = m1_dt$max_runs - m1_dt$value_pb + 
  sapply(m1_dt$outs, function(o) switch(o + 1, m1_vl["['0,0,0,0']"], 
                                        m1_vl["['1,0,0,0']"],
                                        m1_vl["['2,0,0,0']"]))

# Calculate value_created
m1_dt$value_created = m1_dt$runs_scored + m1_dt$value_pa - m1_dt$value_pb
# Calculate BER
m1_dt$BER = m1_dt$value_created / m1_dt$value_max

```


# Model 2 Version 1 BER: position_before and league (no interaction)
```{r}
# Create a value list. Basically a dictionary for faster reference.
state = c(sort(unique(dt[league == "AL"]$PositionBefore_League)), 
          sort(unique(dt[league == "NL"]$PositionBefore_League)))
m2_v1_vl = c(m2_v1_vt$value[1:24], m2_v1_vt$value[1:24] + m2_v1_vt$value[25])
names(m2_v1_vl) = state
m2_v1_vl = c(m2_v1_vl, "end inningAL" = 0, "end inningNL" = 0)

# Create a condensed data.table
m2_v1_dt = dt[, c("res_batter", "PositionBefore_League", "PositionAfter_League",
                  "runs_scored", "max_runs", "runs_scored_EOI", 
                  "outs", "league")]

# Calculate values of position_before and position_after
m2_v1_dt$value_pb = sapply(m2_v1_dt$PositionBefore_League, 
                           function(pb) m2_v1_vl[pb])

# Calculate value of position_after
m2_v1_dt$value_pa = sapply(m2_v1_dt$PositionAfter_League, 
                           function(pa) m2_v1_vl[pa])


# Calculate value_max
m2_v1_dt$value_max = m2_v1_dt$max_runs - m2_v1_dt$value_pb + 
  mapply(function(o, l) ifelse(l == "NL", 
                               switch(o + 1, m2_v1_vl["['0,0,0,0']NL"], 
                                      m2_v1_vl["['1,0,0,0']NL"],
                                      m2_v1_vl["['2,0,0,0']NL"]),
                               switch(o + 1, m2_v1_vl["['0,0,0,0']AL"], 
                                      m2_v1_vl["['1,0,0,0']AL"],
                                      m2_v1_vl["['2,0,0,0']AL"])),
         m2_v1_dt$outs, 
         m2_v1_dt$league)

# Calculate value_created
m2_v1_dt$value_created = m2_v1_dt$runs_scored + m2_v1_dt$value_pa - m2_v1_dt$value_pb
# Calculate BER
m2_v1_dt$BER = m2_v1_dt$value_created / m2_v1_dt$value_max

```
# Model 2 Version 2 BER: position_before and league (include main league effect)
I realize here that since there are only 24 states, Version 2 and Version 3 of Model 2 are essentially the same. And since v3 has a more straight forward interpretation, we will skip down to it. 

# Model 2 Version 3 BER: position_before and league (without main league effect interaction)
```{r}
# Create a value list. Basically a dictionary for faster reference.
state = sort(unique(dt$PositionBefore_League))
m2_v3_vl = m2_v3_vt$value
names(m2_v3_vl) = state
m2_v3_vl = c(m2_v3_vl, "end inningNL" = 0, "end inningAL" = 0)

# Create a condensed data.table
m2_v3_dt = dt[, c("res_batter", "PositionBefore_League", "PositionAfter_League",
                  "runs_scored", "max_runs", "runs_scored_EOI", 
                  "outs", "league")]

# Calculate values of position_before and position_after
m2_v3_dt$value_pb = sapply(m2_v3_dt$PositionBefore_League, 
                           function(pb) m2_v3_vl[pb])

# Calculate value of position_after
m2_v3_dt$value_pa = sapply(m2_v3_dt$PositionAfter_League, 
                           function(pa) m2_v3_vl[pa])


# Calculate value_max
m2_v3_dt$value_max = m2_v3_dt$max_runs - m2_v3_dt$value_pb + 
  mapply(function(o, l) ifelse(l == "NL", 
                               switch(o + 1, m2_v3_vl["['0,0,0,0']NL"], 
                                      m2_v3_vl["['1,0,0,0']NL"],
                                      m2_v3_vl["['2,0,0,0']NL"]),
                               switch(o + 1, m2_v3_vl["['0,0,0,0']AL"], 
                                      m2_v3_vl["['1,0,0,0']AL"],
                                      m2_v3_vl["['2,0,0,0']AL"])),
         m2_v3_dt$outs, 
         m2_v3_dt$league)


# Calculate value_created
m2_v3_dt$value_created = m2_v3_dt$runs_scored + m2_v3_dt$value_pa - m2_v3_dt$value_pb
# Calculate BER
m2_v3_dt$BER = m2_v3_dt$value_created / m2_v3_dt$value_max

```



# Plot BER for players
```{r}
# Load player names
player_names = fread(paste(folder_name, "/Dictionary for Player Names.csv", sep=""), 
                     col.names = c("res_batter", "player_name"))[order(res_batter)]
player_list = sort(unique(dt$res_batter))

# Plate appearance counts
PA_count_table = count(dt, res_batter)

# List of Models of evaluate
model_list = list(m1_dt, m2_v1_dt, m2_v3_dt)

# Create a list BERs for player from our model_list
BER_list = list()
RE24_list = list()
ggplot_list = list()
for (i in 1:length(model_list)) {
  # BER
  BER_list[[i]] = model_list[[i]] %>% 
    group_by(res_batter) %>% 
    summarize(BER_mean = mean(BER))
  # RE24
  RE24_list[[i]] = model_list[[i]] %>% 
    group_by(res_batter) %>% 
    summarize(RE24 = sum(value_created))
  


  
}

# Aggregate in a table
player_table = data.table(res_batter = player_names[res_batter %in%
                                                   player_list]$res_batter,
                          player_names = player_names[res_batter %in%
                                                     player_list]$player_name,
                          BER_m1 = BER_list[[1]]$BER_mean,
                          BER_m2_v1 = BER_list[[2]]$BER_mean,
                          BER_m2_v3 = BER_list[[3]]$BER_mean,
                          RE24_m1 = RE24_list[[1]]$RE24,
                          RE24_m2_v1 = RE24_list[[2]]$RE24,
                          RE24_m2_v3 = RE24_list[[3]]$RE24,
                          PA_count = PA_count_table$n)
```

# Graphs for the models

```{r}
# Histogram of BER for all players
hist_m1.1 = ggplot(player_table, aes(x=BER_m1)) +
  geom_histogram(binwidth = .005) +
  ggtitle("Histogram of BER (All PA)[Basic Model]")
   
  
# Histogram of BER for players with > 100 PA
hist_m1.2 = ggplot(player_table[PA_count > 100], aes(x=BER_m1)) + 
  geom_histogram(binwidth = .005) +
  ggtitle("Histogram of BER (PA > 100)[Basic Model]") 

# Plot BER vs RE24 (no labels)
scatter_m1.1 = ggplot(player_table[PA_count >100], 
                    aes(x = BER_m1, y = RE24_m1, label = player_names)) +
  geom_point() + 
  geom_smooth() + 
  geom_smooth(method="lm", color = "red") + 
  xlab("BER") +
  ggtitle("RE24 vs BER (PA > 100)[Basic Model]")

# Plot BER vs RE24 (labels)
scatter_m1.2 = ggplot(player_table[PA_count >100], 
                      aes(x = BER_m1, y = RE24_m1, label = player_names)) +
  geom_point() + 
  geom_smooth() + 
  geom_text(aes(label = player_names)) +
  geom_smooth(method="lm", color = "red") + 
  xlab("BER") + 
  ggtitle("RE24 vs BER (PA > 100)[Basic Model]")


# Model 2 v1 ########################################
hist_m2_v1.1 = ggplot(player_table, aes(x=BER_m2_v1)) +
  geom_histogram(binwidth = .005) +
  ggtitle("Histogram of BER (All PA)[League Model]")
  

hist_m2_v1.2 = ggplot(player_table[PA_count > 100], aes(x=BER_m2_v1)) +
  geom_histogram(binwidth = .005) + 
  ggtitle("Histogram of BER (PA > 100)[League Model]")

# Plot BER vs RE24 (no labels)
scatter_m2_v1.1 = ggplot(player_table[PA_count >100], 
                    aes(x = BER_m2_v1, y = RE24_m2_v1, label = player_names)) +
  geom_point() + 
  geom_smooth() + 
  geom_smooth(method="lm", color = "red") + 
  xlab("BER") +
  ggtitle("RE24 vs BER (PA > 100)[Basic Model]")

# Plot BER vs RE24 (labels)
scatter_m2_v1.2 = ggplot(player_table[PA_count >100], 
                      aes(x = BER_m2_v1, y = RE24_m2_v1, label = player_names)) +
  geom_point() + 
  geom_smooth() + 
  geom_text(aes(label = player_names)) +
  geom_smooth(method="lm", color = "red") + 
  xlab("BER") + 
  ggtitle("RE24 vs BER (PA > 100)[Basic Model]")

# Model 2 v3 ##########################################
hist_m2_v3.1 = ggplot(player_table, aes(x=BER_m2_v3)) +
  geom_histogram(binwidth = .005) +
  ggtitle("Histogram of BER (All PA)[League Interaction Model]")
  

hist_m2_v3.2 = ggplot(player_table[PA_count > 100], aes(x=BER_m2_v3)) +
  geom_histogram(binwidth = .005) + 
  ggtitle("Histogram of BER (PA > 100)[League Interaction Model]")


scatter_m2_v3.1 = ggplot(player_table[PA_count >100], 
                    aes(x = BER_m2_v3, y = RE24_m2_v3, label = player_names)) +
  geom_point() + 
  geom_smooth() + 
  geom_smooth(method="lm", color = "red") + 
  xlab("BER") +
  ggtitle("RE24 vs BER (PA > 100)[League Interaction Model]")

scatter_m2_v3.2 = ggplot(player_table[PA_count >100], 
                      aes(x = BER_m2_v3, y = RE24_m2_v3, label = player_names)) +
  geom_point() + 
  geom_smooth() + 
  geom_text(aes(label = player_names)) +
  geom_smooth(method="lm", color = "red") + 
  xlab("BER") + 
  ggtitle("RE24 vs BER (PA > 100)[League Interaction Model]")

grid.arrange(hist_m1.1, hist_m1.2, scatter_m1.2,
             hist_m2_v1.1, hist_m2_v1.2, scatter_m2_v1.2,
             hist_m2_v3.1, hist_m2_v3.2, scatter_m2_v3.2,
             ncol = 3)


# Split NL and AL players ###################################

player_leagues = dt[, c("res_batter", "league")]


```










# Extraneous code for plotting and exploration (Not included in output)

```{r, eval=F}
# Set the data.table for the model we want to examine
model_dt = m2_v3_dt

player_names = fread("/Users/txl/Google Drive/Sports Analytics/Final Data/Dictionary for Player Names.csv", col.names = c("res_batter", "player_name"))[order(res_batter)]
player_list = sort(unique(model_dt$res_batter))
BER_mean_table = model_dt %>% group_by(res_batter) %>% summarize(BER_mean = mean(BER))
RE24 = model_dt %>% group_by(res_batter) %>% summarize(RE24 = sum(value_created))
PA_count_table = count(model_dt, res_batter)

# Create a BER table for players
BER_table = data.table(res_batter = player_names[res_batter %in%
                                                   player_list]$res_batter,
                       player_names = player_names[res_batter %in%
                                                     player_list]$player_name,
                       BER_mean = BER_mean_table$BER_mean, 
                       RE24,
                       PA_count = PA_count_table$n)

BER_table[, res_batter := NULL]

# Top 10 Player BERs with PA > 100
head(BER_table[PA_count > 100][order(BER_mean, decreasing = T)][, -c("res_batter")], 10)
# Bottom 10 
head(BER_table[PA_count > 100][order(BER_mean, decreasing = F)][,-c("res_batter")], 10)


# Top 10 Player BERs 
head(BER_table[order(RE24, decreasing = T)][, -c("res_batter")], 10)
# Bottom 10 
head(BER_table[order(RE24, decreasing = F)][,-c("res_batter")], 10)

# Histogram of AB_count
ggplot(BER_table, aes(x=PA_count)) + 
  geom_histogram(binwidth = 10) + 
  ggtitle("Histogram of PA_count")


# Histogram of BER 
ggplot(BER_table, aes(x=BER_mean)) +
  ggtitle("Histogram of BER_mean (All PA)") +
  geom_histogram(binwidth = .005)

# Histogram of BER for players with > 100 PA
ggplot(BER_table[PA_count > 100], aes(x=BER_mean)) + 
  ggtitle("Histogram of BER_mean (PA > 100)") + 
  geom_histogram(binwidth = .005)

# Histogram of BER for players with > 500 PA
ggplot(BER_table[PA_count > 500], aes(x=BER_mean)) + 
  ggtitle("Histogram of BER_mean (PA > 500)") + 
  geom_histogram(binwidth = .005)

# Plot BER vs RE24 (no labels)
ggplot(BER_table[,-c("res_batter")][PA_count >100], aes(x = BER_mean, y = RE24, label = player_names)) +
  geom_point() + 
  geom_smooth() + 
  geom_smooth(method="lm", color = "red") + 
  xlab("BER") +
  ggtitle("RE24 vs BER (PA > 100)")

# Plot BER vs RE24 (labels)
ggplot(BER_table[,-c("res_batter")][PA_count >100], aes(x = BER_mean, y = RE24, label = player_names)) +
  geom_point() + 
  geom_smooth() + 
  geom_text(aes(label = player_names)) +
  geom_smooth(method="lm", color = "red") + 
  xlab("BER") + 
  ggtitle("RE24 vs BER (PA > 100)")

```