---
title: "BER_tlian_v3"
author: "TXL"
date: "5/24/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(gridExtra)

# Source tool functions file
source("BER_tool_functions.R")




```


# BER Version 3

We use the following methodology to calculate BERs for a given year.
1. For year $Y$, aggregate all data in years $[Y-2, Y+2]$ to use in regression model. Calculate across time.
2. Create a BER table for players for each season
Note: We have data for years 1975-2018. Easy to start at year 1977, but need to decide what to do with year 2018.

# Aggregate data
We write a for-loop covering only one year as it will be easily translatable across time. 
```{r}

export_folder = "ExportCSV"

years = 2016
for (y in years) {
  print(paste("Aggregating data for year:", as.character(y)))
  datalist = list()
  for (i in (y-2):(y+2)) {
    if (i == y) {
      current_year_dt = import_season_data(i)
      datalist[[i]] = current_year_dt
    }
    else {
      datalist[[i]] = import_season_data(i)
    }

  }
  dt = do.call(rbind, datalist) # Aggregate data.table
  
  # Now we run the regression model with this data.table
  # m$reg contains the full linear model
  # m$reg_vl contains the list of values for each state
  m = calculate_BER_league(dt)
  
  BER_dt = calculate_BER(current_year_dt, m$vl)
  
  player_table = create_player_table(BER_dt)
  

  # Export CSVs
  # Value list
  print("Exporting value_list...")
  write.csv(pt, paste0("valuelist_", as.character(y), ".csv"))
  
  print("Exporting BER_dt...")
  write.csv(BER_dt, paste0("BER_dt_", as.character(y), ".csv"))

  print("Exporting player_table...")
  write.csv(player_table, paste0("player_table_", as.character(y), ".csv"))
 
   print("Successfully exported all tables. ")
  

  
  
  
  
  
  
  
  
  
  
}




```



